---
output: 
  github_document:
    toc: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```


# Runner

[![Travis-CI Build Status](https://travis-ci.org/elo2zero/runner.svg?branch=master)](https://travis-ci.org/elo2zero/runner)
[![MIT License](https://badges.frapsoft.com/os/mit/mit.svg)](https://opensource.org/licenses/mit-license.php)
[![Coverage status](https://codecov.io/gh/elo2zero/runner/branch/master/graph/badge.svg)](https://codecov.io/github/elo2zero/runner?branch=master)


Running functions for R vector writen in Rcpp

## Installation

You can install runner from github with:

```{r gh-installation, eval=FALSE, include=FALSE}
# install.packages("devtools")
devtools::install_github("elo2zero/runner")
```

## Examples
The main idea of the package is to provide running operations on R vectors. Running functions are these which are applied to all elements up to actual one. Typical example implemented already in `base` are `cumsum`, `cummean`, `cummin` etc. Functions provided in this package works similar but with extended functionality such as handling `NA` and custom window size. Most functions provided in package are basing on the same logic
1. `k` window size denotes number of elements from i-th backwards, where functions are calculated.  
(obrazek pokazujący ruchome okno)
2. `na_rm` handling missing equivalent to `na.rm`. In case of running functions, `NA` is replaced with last finite value.
3. `na_pad` if window size exceeds number of available elements, than first `k-1` are filled with `NA`.
4. `which` In case of running index  

### window_run
Function creates list of windows. Because `runner` doesn't provide limited functionality, one can calculate own running-window-statistics oneself. 

```{r run_window, echo=FALSE}
library(runner)
library(magrittr)
set.seed(11)
x1    <- 1:5
k     <- c(1,2,3,3,2)

window_run( x1, k = k )
```

Such windows can be used in further calculations, with any R function.

```{r run_window2, echo=FALSE} 
window_run( x1, k = k ) %>%
  lapply(sum) %>%
  unlist
```

### min_run and max_run

```{r min_run_example}
library(ggplot2)
set.seed(11)
x1 <- rnorm( 30 ) %>% cumsum
x2 <- sample( c( rep(NA,10), rnorm(20)) ) %>% cumsum
k  <- sample( 1:4, 30, replace=TRUE) %>% cumsum

min1 <- min_run(x1)
min2 <- min_run(x2, na_rm = TRUE)
min3 <- min_run(x2, na_rm = TRUE, k=4)
min4 <- min_run(x2, na_rm = FALSE, k=k)

max1 <- max_run(x1)
max2 <- max_run(x2, na_rm = TRUE)
max3 <- max_run(x2, na_rm = TRUE, k=4)
max4 <- max_run(x2, na_rm = FALSE, k=k)

data.frame( x1, min1, i = 1:30) %>%
  ggplot(aes(x=i, y=x1)) +
  geom_line() +
  geom_point(aes(y = min1), color="red") +
  geom_point(aes(y = max1), color="blue") +
  ggtitle("running minimum")

```
