---
title: "Built-in functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Built-in functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

While `runner::runner` can apply any R function, built-in functions provide
C++ implementations of common operations that are orders of magnitude faster.
All built-in functions accept the same `k`, `lag`, `idx`, `at`, and `na_pad`
arguments as `runner`. See `vignette("getting_started_with_runner")` for an
introduction to these parameters.

## Running aggregations

### `min_run`, `max_run`, `sum_run`, `mean_run`

These compute running minimum, maximum, sum, and mean respectively. By default
(`na_rm = TRUE`), `NA` values are ignored. The default window is cumulative
(equivalent to `cummin`, `cummax`, etc.), but `k` restricts it to a fixed
size.

The diagram below shows `min_run` at i=8 with three configurations:

```
              running minimum at i=8
                                  |
x:       1  -5   1  -3  NA  NA  NA   1  -1  NA  -2   3
                                  v
default: [--------window---------]-5  (cumulative, all elements 1..8)
k=5:              [----window----]-3  (elements 4..8)
na_rm=F: [--------window---------]NA  (NA present -> result is NA)
```

```{r min_run_example, echo=TRUE}
library(runner)
x <- c(1, -5, 1, -3, NA, NA, NA, 1, -1, NA, -2, 3)

data.frame(
  x,
  default = min_run(x, na_rm = TRUE),
  k_5     = min_run(x, k = 5, na_rm = TRUE),
  narm_f  = min_run(x, na_rm = FALSE)
)
```

### Using `idx` for time-based windows

When data points are not equally spaced, set `idx` so that `k` refers to
index distance rather than element count. In the example below, a 5-day sum
window spans different numbers of elements depending on date spacing:

```
     x        idx         sum_run(k=5)
 -0.5910   1970-01-03     -0.5910
  0.0266   1970-01-06     -0.5644
 -1.5166   1970-01-09     -1.4900
 -1.3627   1970-01-12     -2.8792
  1.1785   1970-01-13     -1.7007
 -0.9342   1970-01-16     -1.1183  <- 5 days [13..17]: rows 5,6,7
  1.3236   1970-01-17      1.5679  <- 5 days [13..17]: rows 5,6,7
  0.6249   1970-01-19      1.0144  <- 5 days [15..19]: rows 6,7,8
```

```{r idx_run_example, echo=TRUE}
x <- c(-0.5910, 0.0266, -1.5166, -1.3627, 1.1785, -0.9342, 1.3236, 0.6249)
idx <- as.Date(c(
  "1970-01-03", "1970-01-06", "1970-01-09", "1970-01-12",
  "1970-01-13", "1970-01-16", "1970-01-17", "1970-01-19"
))

sum_run(x, k = 5, idx = idx)
```

The `lag` argument shifts the window by index distance (or by element count
if `idx` is not set):

```{r lag_run_example, echo=TRUE}
sum_run(x, k = 5, lag = 2, idx = idx)
```

### `streak_run`

Counts consecutive identical values within each window. The `na_rm` argument
controls how `NA` is handled — when `na_rm = FALSE`, `NA` resets the streak;
when `na_rm = TRUE`, `NA` is skipped and the streak continues through it.

```
              window streaks at i=9
                                    |
x:      A   B   A   A   B   B   B  NA   B   B   A   B
                                    v
full:   1   1   1   2   1   2   3   3   4  (NA skipped, streak continues)
k=4:                        1   2       1  (window of 4, na_rm=F: NA resets)
na_rm=T:                    1   2   2   3  (window of 4, na_rm=T: NA skipped)
```

```{r}
x <- c("A", "B", "A", "A", "B", "B", "B", NA, "B", "B", "A", "B")
data.frame(
  x,
  s0 = streak_run(x),
  s1 = streak_run(x, k = 4, na_rm = FALSE),
  s2 = streak_run(x, k = 4)
)
```

Streaks work with `idx` too — for example, counting consecutive wins in a
5-day period:

```{r}
x <- c("W", "W", "L", "L", "L", "W", "L", "L")
idx <- as.Date(c(
  "2019-01-03", "2019-01-06", "2019-01-09", "2019-01-12",
  "2019-01-13", "2019-01-16", "2019-01-17", "2019-01-19"
))

data.frame(
  idx, x,
  streak_5d     = streak_run(x, k = 5, idx = idx),
  streak_5d_lag = streak_run(x, k = 5, lag = 1, idx = idx)
)
```

## Utility functions

### `lag_run`

Like `stats::lag`, but supports index-based lag via `idx`. When `idx` is set,
the lag refers to index distance — so `lag = 3` with a date index means "the
value 3 days ago", not "the value 3 positions ago":

```{r}
x <- c(-0.5910, 0.0266, -1.5166, -1.3627, 1.1785, -0.9342, 1.3236, 0.6249)
idx <- as.Date(c(
  "1970-01-03", "1970-01-06", "1970-01-09", "1970-01-12",
  "1970-01-13", "1970-01-16", "1970-01-17", "1970-01-19"
))

lag_run(x, lag = 3, idx = idx)
```

### `fill_run`

Replaces `NA` values with the preceding non-NA value. Two options modify the
default behavior:

- `run_for_first = TRUE` — fills leading `NA`s (at the start of the vector)
  with the first non-NA value found after them.
- `only_within = TRUE` — only fills `NA`s that are surrounded by the same
  value on both sides (e.g. `"a", NA, "a"` becomes `"a", "a", "a"`, but
  `"a", NA, "b"` stays unchanged).

```
replace NA by preceding value

x:       NA  NA   b   b   a  NA  NA   a   b  NA   a   b
default: NA  NA   b   b   a   a   a   a   b   b   a   b
run:      b   b   b   b   a   a   a   a   b   b   a   b
within:  NA  NA   b   b   a   a   a   a   b  NA   a   b
```

```{r}
x <- c(NA, NA, "b", "b", "a", NA, NA, "a", "b", NA, "a", "b")
data.frame(x,
  f1 = fill_run(x),
  f2 = fill_run(x, run_for_first = TRUE),
  f3 = fill_run(x, only_within = TRUE)
)
```

### `which_run`

Returns the index of the first or last `TRUE` value in each window, controlled
by `which = "first"` or `which = "last"`. When `na_rm = FALSE`, an `NA` in the
window before a `TRUE` makes the result `NA` (because the missing value could
have been `TRUE`).

```
              running which at i=9
                                  |
index:   1   2   3   4   5   6   7   8   9  10  11  12
x:       T   T   T   F  NA   T   F  NA   T   F   T   F
                                  v
default: 1                              (first T in cumulative window)
na_rm=F:              6                 9  (first T in k=5; NA before T -> NA)
na_rm=T:              6                 9  (first T in k=5; NA skipped)
```

```{r}
x <- c(T, T, T, F, NA, T, F, NA, T, F, T, F)
data.frame(
  x,
  s0 = which_run(x, which = "first"),
  s1 = which_run(x, na_rm = FALSE, k = 5, which = "first"),
  s2 = which_run(x, k = 5, which = "last")
)
```
